### 2017-11-29 talk.Rmd
finish first draft of slides

---
title: "Interpretable Machine Learning"
author: "Christoph Molnar"
date: "December 4, 2017"
output:
  ioslides_presentation:
    css: styles.css
---

##  {.emphasizedabit}
Scenario: A company approaches you to predict data scientist salaries with machine learning.

##

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics("images/wow.jpg")
```

# Let's predict data scientist salaries

## What is Machine Learning

Machine learning is a method for teaching computers to make and improve predictions or behaviours based on data.

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics("images/magic.jpg")
```

## Step 1: Find some data
Kaggle conducted an industry-wide survey of data scientists. 
https://www.kaggle.com/kaggle/kaggle-survey-2017

Information asked:

- Compensation
- Demographics
- Job title
- Experience
- ...

```{r load-data, warning=FALSE, include=FALSE}
library('mlr')
library('ggplot2')
library('tidyr')
library('lime')
source('code/prepare-kaggle-data.R')
```

## Step 2: Throw ML on your data
```{r learn, warning=FALSE}
set.seed(42)
task = makeRegrTask(data = survey.dat, target = 'CompensationAmount')
lrn = makeLearner('regr.randomForest', importance=TRUE)
mod = train(lrn, task)
```
## Step 3: Profit. We are done! {.center}

```{r, echo=FALSE, out.width='70%', fig.align='center'}
knitr::include_graphics("images/done-here.gif")
```

##  {.emphasizedabit}

"There is a problem with the model!"

## What problem?

```{r, echo=FALSE, out.width='100%', fig.align='center'}
knitr::include_graphics("images/Hide-the-pain-harold-phone.jpg")
```

## "The older the applicants, the higher the predicted salary, regardless of skills."
```{r, echo=FALSE, out.width='100%', fig.align='center'}
knitr::include_graphics("images/age.jpeg")
```

# Individual Conditional Expectation

##

```{r}
ice = generatePartialDependenceData(mod, task, features ='Age', 
                                    individual = TRUE)
plotPartialDependence(ice) + scale_y_continuous(limits=c(0, NA))
```

##

```{r}
ice.c = generatePartialDependenceData(mod, task, features ='Age', 
          individual = TRUE, center = list(Age=20))
plotPartialDependence(ice.c)
```

# Partial dependence plots

##

```{r}
pdp = generatePartialDependenceData(mod, task, features =c('Age'))
plotPartialDependence(pdp) + scale_y_continuous(limits=c(0, NA))
```

##  {.emphasizedabit}

"We want to understand the model better!"

# Permutation feature importance

##

```{r, warning=FALSE}
feat.imp = getFeatureImportance(mod, type=1)$res
dat = gather(feat.imp, key='Feature', value='Importance') %>% arrange(Importance)
dat$Feature = factor(dat$Feature, levels = dat$Feature)
ggplot(dat)  + geom_point(aes(y=Feature, x = Importance))
```

## Gender?!
```{r, echo=FALSE, fig.align='center', out.width='80%'}
knitr::include_graphics("images/big-mistake.png")
```

##

```{r}
pdp = generatePartialDependenceData(mod, task, features =c('Gender'))
ggplot(pdp$data) + geom_point(aes(x=Gender, y=CompensationAmount)) + 
  geom_segment(aes(x=Gender, xend=Gender, yend=CompensationAmount), y=0) + 
  scale_y_continuous(limits=c(0, NA)) + 
  theme(axis.text.x = element_text(angle = 10, hjust = 1))
```

##

```{r, echo=FALSE, fig.align='center', out.width='100%'}
knitr::include_graphics("images/facepalm.gif")
```

##

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics("images/hidden-pain-bias.jpg")
```

##

```{r, echo=FALSE, fig.align='center', out.width='40%'}
knitr::include_graphics("images/angry.gif")
```

# LIME

## 
```{r, echo=FALSE, fig.align='center'}
set.seed(44)
```
```{r, echo=TRUE, fig.align='center'}
dat = getTaskData(task)
explanation <- lime(dat, mod)
# Explain new instance
explainer <- lime::explain(dat[3, ], explanation, n_features = 3)
plot_features(explainer, ncol=1)
```

## Interested in learning more?

Read my book about "Interpretable Machine Learning"
https://christophm.github.io/xai-book/

```{r, echo=FALSE, fig.align='center', out.width='80%'}
knitr::include_graphics("images/book.png")
```