### 2018-05-29 talk-40min.Rmd
continues talk

---
title: "Interpretable Machine Learning"
author: "Christoph Molnar"
date: "December 4, 2017"
output: ioslides_presentation
css: styles.css
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
  warning = FALSE,
  message = FALSE, 
  fig.align='center')
```

##  {.emphasizedabit}
A client wants you to predict data scientist salaries with machine learning.

##

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics("images/wow.jpg")
```

# Let's predict data scientist salaries

## What is Machine Learning

Machine learning is a method for teaching computers to make and improve predictions or behaviours based on data.

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics("images/magic.jpg")
```

## Step 1: Find some data
Kaggle conducted an industry-wide survey of data scientists. 
https://www.kaggle.com/kaggle/kaggle-survey-2017

Information asked:

- Compensation
- Demographics
- Job title
- Experience
- ...

<font size="2">Contains information from Kaggle ML and Data Science Survey, 2017, which is made available here under the Open Database License (ODbL).</font>

```{r load-data,include=FALSE}
library('mlr')
library('ggplot2')
library('tidyr')
library('lime')
source('code/prepare-kaggle-data.R')
```

## Step 2: Throw ML on your data
```{r learn}
library('mlr')
set.seed(42)
task = makeRegrTask(data = survey.dat, target = 'CompensationAmount')
lrn = makeLearner('regr.randomForest')
mod = train(lrn, task)
```
## Step 3: Profit. {.center}

```{r, echo=FALSE, out.width='70%'}
knitr::include_graphics("images/done-here.gif")
```

##  {.emphasizedabit}

"There is a problem with the model!"

## What problem?

```{r, echo=FALSE}
knitr::include_graphics("images/Hide-the-pain-harold-phone.jpg")
```

## "The older the applicants, the higher the predicted salary, regardless of skills."
```{r, echo=FALSE, out.width='100%'}
knitr::include_graphics("images/age.jpeg")
```

# Partial Dependence Plots

##

```{r}
library("iml")
X = survey.dat[-which(names(survey.dat) == "CompensationAmount")]
predictor = Predictor$new(mod, data = X, y = getTaskTargets(task))
pd = Partial$new(predictor, feature ='Age')
pd$plot() 
```

<font size="2">Goldstein, A., Kapelner, A., Bleich, J., & Pitkin, E. (2013). Peeking Inside the Black Box: Visualizing Statistical Learning with Plots of Individual Conditional Expectation, 1–22. https://doi.org/10.1080/10618600.2014.907095 </font>

<font size="2">Friedman, J. H. (1999). Greedy Function Approximation : A Gradient Boosting Machine. North, 1(3), 1–10. https://doi.org/10.2307/2699986 </font>

##  {.emphasizedabit}

Client: "We want to understand the model better!"

# Permutation feature importance

##

```{r, warning=FALSE}
feat.imp = FeatureImp$new(predictor, loss = 'mae')
feat.imp$plot()
```

<font size="2">Breiman, Leo. "Random forests." Machine learning 45.1 (2001): 5-32. </font>

## Gender?!
```{r, echo=FALSE, out.width='80%'}
knitr::include_graphics("images/big-mistake.png")
```

##

```{r}
pdp = Partial$new(predictor, feature="Gender")
pdp$plot() 
```

```{r}
pdp$results[pdp$results$.type == "pdp", c("Gender", ".y.hat")]
```

##

```{r, echo=FALSE}
knitr::include_graphics("images/hidden-pain-bias.jpg")
```

##

```{r, echo=FALSE, out.width='40%'}
knitr::include_graphics("images/angry.gif")
```

# LIME

## 
```{r, echo=FALSE}
set.seed(44)
```

```{r}
explanation = LocalModel$new(predictor)
explanation$explain(X[15, ])
explanation$plot()
```

<font size="2">Ribeiro, M. T., Singh, S., & Guestrin, C. (2016). “Why Should I Trust You?”: Explaining the Predictions of Any Classifier. Retrieved from http://arxiv.org/abs/1602.04938</font>

# Shapley

## 
```{r}
set.seed(44)
```

```{r}
explanation = Shapley$new(predictor)
explanation$explain(X[15, ])
explanation$plot()
```

# Some Theory

## What is interpretability?

Interpretability is the degree to which a human can understand the cause of a decision. 
But we don't really have a good way to measure that really. Not as easy as benchmarking ML algorithms.

Some things:
sparsity, contrastiveness, social, focus on the abnormal, truthful, coherent with prior beliefs, general and probable.

<font size="2">Miller, Tim. 2017. “Explanation in Artificial Intelligence: Insights from the Social Sciences.” arXiv Preprint arXiv:1706.07269.</font>

## When do we need interpretability?

We need interpretability when the loss function does not cover all constraints.

(like, everything)

## When don't we need interpretability?

When we can capture everything in the loss function and the data collection.
only causal relationships.
perfect operationalization of features.

## What tools do we have?

- Intrinsically interpretable models
- Tools for analyzing specific black box models
- Model-agnostic tools: analysing any black box model
- Interpretability through making data explainable

### Interpretable Models

TODO: Graphic with schematic interpretable model

### Interpretable Models

TODO: Graphic with tree+if-then-rules+decision tree

## Model-agnostic methods
TODO: Graphic with schematic Model-agnostic methods

## Model-agnostic methods

- Partial dependence plots
- LIME
- ...

## Model-specific methods
TODO: Graphic with schematic Model-specific methods

TODO: add images from papers

## Model-specific methods

- RandomForestExplainer
- Visualizing activations of neural networks
-

## Interested in learning more?

Read my book about "Interpretable Machine Learning"
https://christophm.github.io/interpretable-ml-book/

```{r, echo=FALSE, fig.align='center', out.width='30%'}
knitr::include_graphics("images/book.jpg")
```