### 2019-01-08 2018-appliedr/talk.Rmd
clean up repo

---
title: "Interpretable Machine Learning"
author: "Christoph Molnar"
date: "December 4, 2017"
output:
  ioslides_presentation:
    css: ../styles.css
---

##  {.emphasizedabit}
Scenario: A company approaches you to predict data scientist salaries with machine learning.

##

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics("images/wow.jpg")
```

# Let's predict data scientist salaries

## What is Machine Learning

Machine learning is a method for teaching computers to make and improve predictions or behaviours based on data.

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics("images/magic.jpg")
```

```{r child = "../chunks/story-salaries.Rmd"}
```

## Interested in learning more?

Read my book about "Interpretable Machine Learning"
https://christophm.github.io/interpretable-ml-book/

```{r, echo=FALSE, fig.align='center', out.width='80%'}
knitr::include_graphics("images/book.png")
```

```{r child = "../chunks/book.Rmd"}
```