### 2019-01-08 chunks/story-salaries.Rmd

##  {.emphasizedabit}
A client wants you to predict data scientist salaries with machine learning.

## {.center data-background=../images/wow.jpg data-background-size=contain}

# Let's predict data scientist salaries

# What is Machine Learning?

<div class="notes">
Machine learning is a method for teaching computers to make and improve predictions or behaviours based on data.
</div>

## {data-background=../../images/magic.jpg data-background-size=contain}

# Step 1: Find some data

## Step 1: Find some data

Kaggle conducted an industry-wide survey of data scientists. 
https://www.kaggle.com/kaggle/kaggle-survey-2017

Information asked:

- Compensation
- Demographics
- Job title
- Experience
- ...

<font size="2">Contains information from Kaggle ML and Data Science Survey, 2017, which is made available here under the Open Database License (ODbL).</font>

```{r load-data,include=FALSE}
library('mlr')
library('ggplot2')
library('tidyr')
library('lime')
source('../code/prepare-kaggle-data.R')
```

# Step 2: Throw ML on your data

```{r learn}
library('mlr')
set.seed(42)
task = makeRegrTask(data = survey.dat, target = 'CompensationAmount')
lrn = makeLearner('regr.randomForest')
mod = train(lrn, task)
```

## {.center data-background=../images/comp-dog.gif data-background-size=contain}

<div class="notes">
- Random Forest
- Target: Compensation
- All features that you found in the data
- Train, test, ship
</div>

# Step 3: Profit {.center}

## {data-background=../images/done-here.gif data-background-size=contain}
```{r, echo=FALSE, out.width='80%', include = FALSE}
knitr::include_graphics("../images/done-here.gif" )
```

##  {.emphasizedabit}

Client: "There is a problem with the model!"

##  {.emphasizedabit .center data-background=../images/Hide-the-pain-harold-phone.jpg data-background-size=cover}

"What problem?"

```{r, echo=FALSE, out.height='50%', include=FALSE}
knitr::include_graphics("../images/Hide-the-pain-harold-phone.jpg")
```

## {.emphasizedabit}

Client: "The older the candidates, the higher the predicted salaries."

```{r, echo=FALSE, fig.width=16, include=FALSE}
knitr::include_graphics("../images/age.jpeg")
```

##  {.emphasizedabit .center data-background=../images/black-box.gif data-background-size=cover}

<div class="white">
Looking inside the black box
</div>

# How do the features influence my predictions?

## Partial Dependence Plot

```{r out.width="80%"}
library("iml")
X = survey.dat[-which(names(survey.dat) == "CompensationAmount")]
predictor = Predictor$new(mod, data = X, y = getTaskTargets(task))
pd = Partial$new(predictor, feature ='Age')
pd$plot() 
```

<font size="2">Goldstein, A., Kapelner, A., Bleich, J., & Pitkin, E. (2013). Peeking Inside the Black Box: Visualizing Statistical Learning with Plots of Individual Conditional Expectation, 1–22. https://doi.org/10.1080/10618600.2014.907095 </font>

<font size="2">Friedman, J. H. (1999). Greedy Function Approximation : A Gradient Boosting Machine. North, 1(3), 1–10. https://doi.org/10.2307/2699986 </font>

<div class="notes">
- PDP answers: How does the prediction change when a feature input changes?
- Average of ICE curves
</div>

##  {.emphasizedabit}

Client: "We want to understand the model better!"

# What are the most important features?

## Permutation feature importance

```{r, warning=FALSE}
feat.imp = FeatureImp$new(predictor, loss = 'mae')
feat.imp$plot()
```

<font size="2">Breiman, Leo. "Random forests." Machine learning 45.1 (2001): 5-32. </font>

## Gender?! {.emphasizedabit .center data-background=../images/big-mistake.png data-background-size=contain}

```{r, echo=FALSE, out.width='80%', include=FALSE}
knitr::include_graphics("../images/big-mistake.png")
```

## What's the influence of gender on the prediction?

```{r}
pdp = Partial$new(predictor, feature="Gender", ice = FALSE)
knitr::kable(pdp$results[c("Gender", ".y.hat")])
```

## {.center data-background=../images/hidden-pain-bias.jpg data-background-size=contain}

```{r, echo=FALSE, include = FALSE}
knitr::include_graphics("../images/hidden-pain-bias.jpg")
```

## {.center data-background=../images/angry.gif data-background-size=contain}

```{r, echo=FALSE, out.width='40%', include = FALSE}
knitr::include_graphics("../images/angry.gif")
```

# Explaining individual predictions

# Local Models (LIME)

## {.center data-background=../images/lime-fitting-1.png data-background-size=contain}

## 
```{r, echo=FALSE}
set.seed(44)
```

```{r}
explanation = LocalModel$new(predictor)
explanation$explain(X[15, ])
explanation$plot()
```

<font size="2">Ribeiro, M. T., Singh, S., & Guestrin, C. (2016). “Why Should I Trust You?”: Explaining the Predictions of Any Classifier. Retrieved from http://arxiv.org/abs/1602.04938</font>