### 2019-01-08 chunks/tools.Rmd

# What tools do we have?

<div class="notes">
- Intrinsically interpretable models
- Tools for analyzing specific black box models
- Model-agnostic tools: analysing any black box model
- Interpretability through making data explainable
</div>

## Interpretable Models

```{r, echo = FALSE}
knitr::include_graphics("../images/black-box.png")
```

## Interpretable Models

```{r, echo=FALSE}
knitr::include_graphics("../images/white-box.png")
```

<div class="notes">
- Decision trees
- Decision rules
- Generalized regression models
- ...
</div>

## Intepretable Model: Linear Regression {.center data-background=../images/lm.png data-background-size=contain}

## Intepretable Model: Decision Tree {.center data-background=../images/tree.png data-background-size=contain}

## Interpretable Model: Decision Rules {.emphasizedabit}

IF $90m^2\

## Model-specific methods

```{r, echo=FALSE}
knitr::include_graphics("../images/black-box.png")
```

## Model-specific methods

```{r, echo=FALSE}
knitr::include_graphics("../images/specific-black-box.png")
```

<div class="notes">
- Visualizing activations of neural networks
- Random Forest specific feature importance
- xgboost explainer
- problem: new models, not comparable between models, missing flexibility
</div>

## Model-agnostic methods

```{r, echo=FALSE}
knitr::include_graphics("../images/black-box.png")
```

## Model-agnostic methods

```{r, echo=FALSE}
knitr::include_graphics("../images/agnostic-black-box.png")
```

<div class="notes">
- Partial dependence plots
- LIME
- ...
</div>

## Example-based methods

```{r, echo=FALSE}
knitr::include_graphics("../images/black-box.png")
```

## Example-focused methods
```{r, echo=FALSE}
knitr::include_graphics("../images/data-box.png")
```

<div class="notes">
- Prototypes: What is the prototype of a certain class
- Criticisms: Which data points are unusual for a certain class. 
- Counterfactuals: What do I need to change for an instance to change the prediction?
- Anchors: What features do I have to anchor to keep prediction the same?
</div>