### 2019-01-14 chunks/tools.Rmd

# What tools do we have?

<div class="notes">
- Intrinsically interpretable models
- Tools for analyzing specific black box models
- Model-agnostic tools: analysing any black box model
- Interpretability through making data explainable
</div>

##

TODO: Overview (drawing)

- global vs. local
- interpretable model vs. post-hoc
- model-specific vs. model-agnostic
- data type image vs. text vs. tabular

## Interpretable Models

```{r, echo = FALSE}
knitr::include_graphics("../images/black-box.png")
```

## Interpretable Models

```{r, echo=FALSE}
knitr::include_graphics("../images/white-box.png")
```

<div class="notes">
- Decision trees
- Decision rules
- Generalized regression models
- ...
</div>

## Intepretable Model: Linear Regression {.center data-background=../images/lm.png data-background-size=contain}

## Intepretable Model: Decision Tree {.center data-background=../images/tree.png data-background-size=contain}

## Interpretable Model: Decision Rules {.emphasizedabit}

IF $90m^2\leq \text{size} < 110m^2$  AND location $=$ "good" THEN rent is between 1540 and 1890 EUR

## Model-specific methods

```{r, echo=FALSE}
knitr::include_graphics("../images/black-box.png")
```

## Model-specific methods

```{r, echo=FALSE}
knitr::include_graphics("../images/specific-black-box.png")
```

## Model-specific methods

TODO: Example for CNNs

## Model-specific methods

TODO: Example for text (RNNs and attention?)

<div class="notes">
- Visualizing activations of neural networks
- Random Forest specific feature importance
- xgboost explainer
- problem: new models, not comparable between models, missing flexibility
</div>

## Model-agnostic methods

```{r, echo=FALSE}
knitr::include_graphics("../images/black-box.png")
```

## Model-agnostic methods

```{r, echo=FALSE}
knitr::include_graphics("../images/agnostic-black-box.png")
```

<div class="notes">
- Partial dependence plots
- LIME
- ...
</div>

## Model-agnostic Methods

TODO: Drawing of feature effect

## Model-agnostic methods

TODO: Drawing of importance

## Model-agnostic methods: Global Surrogate {.center data-background=../images/global-surrogate.png data-background-size=contain}

## Model-agnostic methods: Local Surrogate {.center data-background=../images/local-surrogate.png data-background-size=contain}

## Example-focused Methods

TODO: Graphic for counterfactuals

## Example-focused Methods

TODO: Graphic for prototypes