### 2021-07-29 2021-bips-practice-phd-defence/2021-bips-practice-phd-defence.Rmd
structure talk

---
title: "Model-agnostic Interpretable Machine Learning"
author: "Christoph Molnar"
date: "2021-08"
institute: "BIPS bremen"
output: 
  beamer_presentation:
    includes:
      in_header: "header_pagenrs.tex"
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
  warning = FALSE,
  message = FALSE, 
  fig.align='center', 
  out.width = '90%',
  dev = "CairoPNG")

ggplot2::theme_set(ggplot2::theme_bw(base_size = 15))
```

# Some ideas

- Cite pitfalls at many places in the talk
- INtroduction with History: INterweave motivation, and also cite history paper
-

# Outline

- A Brief History of Interpretable Modeling
- A Closer Look at Model-agnostic interpretability
- methods:
  - Feature Effects: PDP, ICE and ALE
  - Feature Importance: PFI
- Overview of Stuff I did in my PhD 
- Measuring Model Complexity
- Conditional Feature Importance Using Subgroups
- Inference of Effects

# Before You Pin Me Down

To get this out of the way, here is how I use "Interpretable Machine Learning":

It's an umbrella term for approaches that analyze or describe prediction mechanisms in models that were learned on data.

# A Brief History of Interpretable Modeling

Interpretable Models with data have a longer history, but the field of interpretable machine learning is still very young as consolidated field.

* 1800s: Linear regression models
* 1960: Rule-based ML (AID, CHAID, CART, association rules, ...)
* 2001: Random Forest
* 2021: Deep Learning Boom kicked off
* 2016: Interpretable ML / Explainable AI boom

While 200 years of history, consolidation as larger field only started in 2016.

TODO: Add citation report and or google trends image

# Why Interpretability?

* The more non-linear and complex a model, the more difficult to understand how it operates
* Trust
* Debugging
* Control 
* ...

# Interpretable Models

* GAM and other linear models
* Trees and decision rules
* More complex models, but interpretation requires more work or post-processing

# Model-agnostic Interpretation

Basic idea: Treat the model as a black box, not allowed to look inside.

Idea very powerful: Allows modularity, works for most models, results are comparable across models

TODO: CITE SIPA paper here

# PDP and ICE

# PFI

TODO: Explain how PFI works

# PFI and PDP similarity

CITE paper GC here

# Difference between model-agnostic and model-specific interpetation

* Analyzing components vs. analyzing behavior
* Components not very powerful, unless they directly describe behavior
* For complex model, components do not directly tell us about behavior of model
* Behavior more comparable

# Summary of Thesis

TODO: Draw image with Ipad

In my thesis I did a lot of consolidation work:

- Book
- R package
- History paper
- Pitfalls paper

What if we embrace black box view of models?

Special focus on old, established tools, and make them fit for right use in practice

Also some more specific papers

* subgroups
* inference
* quantifying

and some contributions to

* MOC
* relative feature importance
* SIPA
* feature importance and PDP (GC)

# Paper: Subgroups

## Problem: Dependent features

Show example with dependent features for PFI

## Soluation: Conditional variants

Show solution

* But: Not meaningful for PDP, or other measure more useful (LOESS)
* But: Conditioning unclear
* But: Intepretation difficult

## Our solution: Permute in subgroups

TODO: Show algorithm

## Some results from our paper

TODO: Show results or application

# Paper: IML Inference

## Problem: Uncertainty quantification

PDP and PFI and most other just descriptive tools
Often practictioner draw confidence intervals
but only based on monte carlo approximation

* But what is even the ground truth?
* And what about the uncertainty in the model fit?
* How could we estimate the uncertainty?

## Solution: Treat as statistical estimator

## How to Compute the variance

## Problem of too low variance estimates

# Paper: Quantifying

TODO: Copy as much as possible from the talk that already exists for this paper

Motivation: want to quantify in a model-agnostic way how complex a model is

## Functional Decomposition

Explain how functional decomposition works

## ALE

Formula:

TODO: For estimation show the images

## The three measures

* Mean effect complexity
* non linearity
* number of features

## Using in benchmarking (multi-objective)

## Thank You for Your Attention